---
title: "2020_10_31_Neurosurgery_Residents_Data_Clean_and_Geocoding"
author: "Bob Winkelman"
date: "10/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load Libraries
```{r warning=F, message=F}
library(readxl)
library(janitor)
library(tictoc)
library(here)
library(sf)
library(leaflet)
library(ggmap)
library(plotly)
library(tidyverse)
```


# Read in Data
```{r}

website_links <- read_excel("2020_08_10_US_Neurosurgery_Residents.xlsx",sheet="Website Links") %>%
  clean_names()

resident_data <- read_excel("2020_10_11_US_Neurosurgery_Residents.xlsx") %>%
  clean_names() %>%
  left_join(website_links,by="program")

```


# Geocode Addresses

## Register Google API Key
```{r}
#register_google(key="{enter your key here}")
register_google(key="AIzaSyBxrfbLN79mqK_KG6D9h8ybAq85GqmW2KY")
```

## Geocode Medical School Addresses
```{r message=F, warning=F}
tic()
unique_medical_school_addresses <- resident_data %>%
  distinct(medical_school,medical_school_address) %>%
  mutate(medical_school = ifelse(is.na(medical_school),"Not Reported",medical_school)) %>%
  mutate(medical_school_address = str_replace_all(medical_school_address,"\r|\n"," ")) %>%
  filter(!is.na(medical_school_address)) %>%
  mutate_geocode(medical_school_address) %>%
  select(-medical_school_address) %>%
  rename(medical_school_lat = lat,
         medical_school_long = lon)
toc()
```


## Geocode Program Addresses
```{r message=F, warning=F}
tic()
unique_program_addresses <- resident_data %>%
  distinct(program,program_address) %>%
  mutate(program_address = str_replace_all(program_address,"\r|\n"," ")) %>%
  filter(!is.na(program_address)) %>%
  mutate_geocode(program_address) %>%
  select(-program_address) %>%
  rename(program_lat=lat,
         program_long=lon)
toc()
```

## Join Geocode Data to Dataset and Calculate Distance from Med School to Programs

```{r}
resident_data_w_geo_code <- resident_data %>% 
  mutate(medical_school = ifelse(is.na(medical_school),"Not Reported",medical_school)) %>%
  left_join(unique_medical_school_addresses,
            by="medical_school") %>%
  left_join(unique_program_addresses,
            by="program") %>%
  mutate(subject_id = row_number(),
         medical_school_long_lat = map2(medical_school_long,
                                        medical_school_lat,
                                        function(x,y) c(x,y)),
         program_long_lat = map2(program_long,
                                 program_lat,
                                        function(x,y) c(x,y)),
         distance_meter = map2_dbl(medical_school_long_lat,
                                   program_long_lat,
                                   function(x,y)
                                     geosphere::distHaversine(x, y)),
         distance_miles = distance_meter/1609.344)
```

# Tidy Data and Add Summary Data

```{r }

final_resident_data_w_geo_code <- resident_data_w_geo_code %>%
  mutate(intern_year = case_when(year=="PGY-7" ~ year_website_updated-6,
                                 year=="PGY-6" ~ year_website_updated-5,
                                 year=="PGY-5" ~ year_website_updated-4,
                                 year=="PGY-4" ~ year_website_updated-3,
                                 year=="PGY-3" ~ year_website_updated-2,
                                 year=="PGY-2" ~ year_website_updated-1,
                                 year=="PGY-1" ~ year_website_updated-0),
         program_update_year = paste0(program," (",year_website_updated,")")) %>%
  group_by(program) %>%
  mutate(avg_miles = round_half_up(mean(distance_miles,na.rm=T),digits=1),
         median_miles = round_half_up(median(distance_miles,na.rm=T),digits=1),
         num_residents = n(),
         num_residents_2 = num_residents,
         num_w_med_school_info = sum(!is.na(medical_school)),
         pct_w_med_school_info = round_half_up(num_w_med_school_info/n()*100,digits=1),
         num_w_missing_med_school_info = num_residents-num_w_med_school_info,
         pct_w_missing_med_school_info=100-pct_w_med_school_info,
         num_home_students = sum(home_student_y_n=="y",na.rm=T),
         pct_home_students = round_half_up(num_home_students/n()*100,digits=1), 
         num_female_residents = sum(m_f=="f",na.rm=T),
         pct_female_residents = round_half_up(num_female_residents/n()*100,digits=1),
         num_male_residents = num_residents - num_female_residents,
         pct_male_residents = 100 - pct_female_residents,
         num_international_student = sum(international_student=="y",na.rm=T),
         pct_international_student = round_half_up(num_international_student/n()*100,digits=1),
         num_advanced_degree = sum(!is.na(additional_degree_s),na.rm=T),
         pct_advanced_degree = round_half_up(num_advanced_degree/n()*100,digits=1),
         num_phd = sum(str_detect(additional_degree_s,"PhD"),na.rm=T),
         pct_phd = round_half_up(num_phd/n()*100,digits=1)) %>%
  ungroup() %>%
  group_by(program,year) %>%
  mutate(year_avg_miles = round_half_up(mean(distance_miles,na.rm=T),digits=1),
         year_median_miles = round_half_up(median(distance_miles,na.rm=T),digits=1)) %>%
  ungroup() %>%
  group_by(medical_school) %>%
  mutate(avg_miles_ms = round_half_up(mean(distance_miles,na.rm=T),digits=1),
         median_miles_ms = round_half_up(median(distance_miles,na.rm=T),digits=1),
         num_residents_ms = n(),
         num_residents_ms_2 = num_residents_ms,
         num_home_students_ms = sum(home_student_y_n=="y",na.rm=T),
         pct_home_students_ms = round_half_up(num_home_students_ms/n()*100,digits=1), 
         num_female_residents_ms = sum(m_f=="f",na.rm=T),
         pct_female_residents_ms = round_half_up(num_female_residents_ms/n()*100,digits=1),
         num_male_residents_ms = num_residents_ms - num_female_residents_ms,
         pct_male_residents_ms = 100 - pct_home_students_ms,
         num_advanced_degree_ms = sum(!is.na(additional_degree_s),na.rm=T),
         pct_advanced_degree_ms = round_half_up(num_advanced_degree_ms/n()*100,digits=1),
         num_phd_ms = sum(str_detect(additional_degree_s,"PhD"),na.rm=T),
         pct_phd_ms = round_half_up(num_phd_ms/n()*100,digits=1)) %>%
  mutate(distance_miles = round_half_up(distance_miles,digits=1)) %>%
  ungroup() %>%
  select(subject_id,everything())
```


```{r}
final_resident_data_w_geo_code %>%
  write_rds(here(path="Data",
                 "2020_10_31_resident_data_w_geo_code.rds"))
```


# Dataset Transformation for App 

## Program Summary Data
```{r}
wide_sum_data_program <- final_resident_data_w_geo_code %>%
    select(program,
           #program_website=website,
           #year_website_updated,
           #date_accessed,
           program_lat,program_long,
           medical_school,medical_school_lat,medical_school_long,
           distance_miles) %>%
    #mutate(program_w_label = paste0('<a href="',program_website,'">',program,"(",year_website_updated,")",'</a>')) %>%
    group_by(program) %>%
    add_count(medical_school,name="num_residents") %>%
    ungroup() %>% 
    distinct(program,program_lat,program_long,
             medical_school,medical_school_lat,medical_school_long,
             distance_miles,num_residents) %>%
    mutate(medical_school_name = medical_school,
           num_residents_2 = num_residents,
           distance_miles = round_half_up(distance_miles)) %>%
    gather(key,value,medical_school_name,num_residents,distance_miles) %>%
    mutate(key = paste0("<b>",key,"</b>")) %>%
    group_by(program,program_lat,program_long,
             medical_school,medical_school_lat,medical_school_long) %>%
    nest(data=c(key,value)) %>%
    ungroup() %>%
    mutate(
        data = map(.x=data, ~.x %>%
                       mutate(key = case_when(str_detect(key,"medical_school_name")~ str_replace(key,"medical_school_name","School Name:"),
                                              str_detect(key,"num_residents")~ str_replace(key,"num_residents","# of Residents:"),
                                              str_detect(key,"distance_miles")~ str_replace(key,"distance_miles","Distance (miles):")))),
        label = map(data, knitr::kable, format="html",escape=FALSE,col.names=c("",""))) %>%
    arrange(program,desc(num_residents_2))
```

```{r}
wide_sum_data_program %>%
  write_rds(here(path="Data",
                 "2020_10_31_wide_sum_data_program.rds"))
```


```{r}
wide_sum_data_med_school <- final_resident_data_w_geo_code %>% 
    select(program,website,year_website_updated,program_lat,program_long,
           medical_school,medical_school_lat,medical_school_long,
           distance_miles) %>%
    mutate(program_w_label = paste0('<a href="',website,'">',program," (",year_website_updated,")",'</a>')) %>%
    group_by(medical_school) %>%
    add_count(program,name="num_residents") %>%
    ungroup() %>% 
    distinct(program,program_w_label,program_lat,program_long,
             medical_school,medical_school_lat,medical_school_long,
             distance_miles,num_residents) %>%
    mutate(program_name = program,
           num_residents_2 = num_residents,
           distance_miles = round_half_up(distance_miles)) %>%
    gather(key,value,program_name,num_residents,distance_miles) %>%
    mutate(key = paste0("<b>",key,"</b>")) %>%
    group_by(program,program_w_label,program_lat,program_long,
             medical_school,medical_school_lat,medical_school_long) %>%
    nest(data=c(key,value)) %>%
    ungroup() %>%
    mutate(
        data = map(.x=data, ~.x %>%
                       mutate(key = case_when(str_detect(key,"program_name")~ str_replace(key,"program_name","Program Name:"),
                                              str_detect(key,"num_residents")~ str_replace(key,"num_residents","# of Residents:"),
                                              str_detect(key,"distance_miles")~ str_replace(key,"distance_miles","Distance (miles):")))),
        label = map(data, knitr::kable, format="html",escape=FALSE,col.names=c("",""))) %>%
    arrange(program,desc(num_residents_2))

```

```{r}
wide_sum_data_med_school %>%
  write_rds(here(path="Data",
                 "2020_10_31_wide_sum_data_med_school.rds"))
```

## Summary Program Long
```{r}
long_sum_data_program <- wide_sum_data_program%>%
    pivot_longer(cols=c(medical_school_long,medical_school_lat,program_long,program_lat),
                 names_to = "location_info",
                 values_to= c("coord")) %>%
    mutate(location_type = str_extract(location_info,"medical_school|program"),
           location_coord_type = str_extract(location_info,"lat|long")) %>%
    select(-location_info) %>%
    pivot_wider(names_from=location_coord_type,
                values_from = coord)
```

```{r}
long_sum_data_program %>% 
  write_rds(here(path="Data",
                 "2020_10_31_long_sum_data_program.rds"))
```



## Data Details Program

```{r}
program_data_details <- final_resident_data_w_geo_code %>%
    distinct(program,
             website,
             year_website_updated,
             avg_miles,median_miles,num_residents,
             num_home_students,pct_home_students,
             num_female_residents,pct_female_residents,
             num_male_residents,pct_male_residents,
             num_international_student,pct_international_student,
             num_advanced_degree,pct_advanced_degree,
             num_phd,pct_phd) %>%
    mutate(program_w_label = paste0('<a href="',website,'">',program," (",year_website_updated,")",'</a>')) %>%
    mutate(data_type="By Program") %>%
    select(-c(website,year_website_updated)) %>%
    select(data_type,
           primary_name=program_w_label,
           primary_name_wo_label=program,
           everything())
```

```{r}
program_data_details %>% 
  write_rds(here(path="Data",
                 "2020_10_31_program_data_details.rds"))

```

## Data Details Medical School
```{r}
medical_school_data_details <- final_resident_data_w_geo_code %>%
    distinct(medical_school,
             avg_miles_ms,median_miles_ms,num_residents_ms,
             num_home_students_ms,pct_home_students_ms,
             num_female_residents_ms,pct_female_residents_ms,
             num_male_residents_ms,pct_male_residents_ms,
             num_advanced_degree_ms,pct_advanced_degree_ms,
             num_phd_ms,pct_phd_ms) %>%
    mutate(medical_school_2 = medical_school) %>%
    mutate(data_type="By Medical School",
           num_international_student_ms=NA,
           pct_international_student_ms=NA) %>%
    select(data_type,
           primary_name=medical_school,
           primary_name_wo_label=medical_school_2,
           avg_miles=avg_miles_ms,
           median_miles=median_miles_ms,
           num_residents=num_residents_ms,
           num_home_students=num_home_students_ms,pct_home_students=pct_home_students_ms,
           num_female_residents=num_female_residents_ms,pct_female_residents=pct_female_residents_ms,
           num_male_residents=num_male_residents_ms,pct_male_residents=pct_male_residents_ms,
           num_international_student=num_international_student_ms,pct_international_student=pct_international_student_ms,
           num_advanced_degree=num_advanced_degree_ms,pct_advanced_degree=pct_advanced_degree_ms,
           num_phd=num_phd_ms,pct_phd=pct_phd_ms)
```

```{r}
medical_school_data_details %>% 
  write_rds(here(path="Data",
                 "2020_10_31_medical_school_data_details.rds"))
```


```{r}
data_w_geo_code_sum_program <- final_resident_data_w_geo_code %>%
    distinct(program,
             website,
             date_accessed,
             avg_miles,median_miles,num_residents,num_w_med_school_info,pct_w_med_school_info,
             num_w_missing_med_school_info,pct_w_missing_med_school_info,num_home_students,
             pct_home_students,num_female_residents,pct_female_residents,num_male_residents,
             num_international_student,
             pct_male_residents,num_advanced_degree,pct_advanced_degree,num_phd,pct_phd) %>%
    mutate(data_type="By Program") %>%
    select(data_type,
           primary_name=program,
           program_website=website,
           date_accessed,
           avg_miles,
           num_residents,
           num_home_students,
           num_female_residents,
           num_male_residents,
           num_international_student,
           num_advanced_degree,
           num_phd) %>%
    mutate(program_w_website_link = paste0('<a href="',program_website,'">',primary_name,'</a>'),
           date_accessed = as.character(format.Date(date_accessed,format="%m/%d/%Y"))) %>% 
    select(-program_website) %>%
    gather(key,value,3:12) %>%
    nest(data=c(key,value)) %>%
    mutate(
        data = map(.x=data, ~.x %>%
                       mutate(key = case_when(str_detect(key,"program_w_website_link")~ str_replace(key,"program_w_website_link","Website Link:"),
                                              str_detect(key,"date_accessed")~ str_replace(key,"date_accessed","Date Accessed:"),
                                              #str_detect(key,"num_residents")~ str_replace(key,"num_residents","# of Residents:"),
                                              str_detect(key,"num_residents")~ str_replace(key,"num_residents","# of Residents:"),
                                              str_detect(key,"avg_miles")~ str_replace(key,"avg_miles","Avg. Dist. (miles):"),
                                              str_detect(key,"num_home_students")~str_replace(key,"num_home_students","# of Home Students:"),
                                              str_detect(key,"num_male_residents")~ str_replace(key,"num_male_residents","# Male Residents:"),
                                              str_detect(key,"num_female_residents")~ str_replace(key,"num_female_residents","# Female Residents:"),
                                              str_detect(key,"num_international_student")~ str_replace(key,"num_international_student","# International Residents:"),
                                              str_detect(key,"num_advanced_degree")~ str_replace(key,"num_advanced_degree","# w/ Advanced Degree:"),
                                              str_detect(key,"num_phd")~ str_replace(key,"num_phd","# w/ PhD:")),
                              key = factor(key, levels= c("Website Link:","Date Accessed:","# of Residents:","Avg. Dist. (miles):","# of Home Students:","# Male Residents:","# Female Residents:","# International Residents:","# w/ Advanced Degree:","# w/ PhD:"))) %>%
                     arrange(key)
                   ),
        label = map(data, knitr::kable, format="html",escape=FALSE,col.names=c("","")))

```

```{r}
data_w_geo_code_sum_program %>%
  write_rds(here(path="Data",
                 "2020_10_31_data_w_geo_code_sum_program.rds"))
```




```{r}
data_w_geo_code_sum_medical_school <- final_resident_data_w_geo_code %>% 
    distinct(medical_school,
             avg_miles_ms,
             num_residents_ms,
             num_home_students_ms,
             num_female_residents_ms,
             num_male_residents_ms,
             num_phd_ms,
             num_advanced_degree_ms) %>%
    mutate(data_type="By Medical School") %>%
    select(data_type,
           primary_name=medical_school,
           avg_miles=avg_miles_ms,
           num_residents=num_residents_ms,
           num_home_students=num_home_students_ms,
           num_female_residents=num_female_residents_ms,
           num_male_residents=num_male_residents_ms,
           num_advanced_degree=num_advanced_degree_ms,
           num_phd=num_phd_ms) %>%
    gather(key,value,3:9) %>%
    distinct() %>%
    nest(data=c(key,value)) %>%
    mutate(
        data = map(.x=data, ~.x %>%
                       mutate(key = case_when(str_detect(key,"num_residents")~ str_replace(key,"num_residents","# of Residents:"),
                                              str_detect(key,"avg_miles")~ str_replace(key,"avg_miles","Avg. Dist. (miles):"),
                                              str_detect(key,"num_home_students")~str_replace(key,"num_home_students","# of Home Students:"),
                                              str_detect(key,"num_male_residents")~ str_replace(key,"num_male_residents","# Male Residents:"),
                                              str_detect(key,"num_female_residents")~ str_replace(key,"num_female_residents","# Female Residents:"),
                                              str_detect(key,"num_international_student")~ str_replace(key,"num_international_student","# International Residents:"),
                                              str_detect(key,"num_advanced_degree")~ str_replace(key,"num_advanced_degree","# w/ Advanced Degree:"),
                                              str_detect(key,"num_phd")~ str_replace(key,"num_phd","# w/ PhD:")
                       ))),
        label = map(data, knitr::kable, format="html",escape=FALSE,col.names=c("","")))

```


```{r}
data_w_geo_code_sum_medical_school %>%
  write_rds(here(path="Data",
                 "2020_10_31_data_w_geo_code_sum_medical_school.rds"))

```



```{r}
overtime_data <- final_resident_data_w_geo_code %>%
  group_by(intern_year) %>%
  summarise(total_n = n(),
            pct_female_residents = round_half_up(sum(m_f == "f",na.rm=T)/total_n*100,digits=1),
            pct_international_student = round_half_up(sum(international_student == "y",na.rm=T)/total_n*100,digits=1),
            pct_advanced_degree = round_half_up(sum(!is.na(additional_degree_s),na.rm=T)/total_n*100,digits=1),
            pct_phd = round_half_up(sum(str_detect(additional_degree_s,"PhD"),na.rm=T)/total_n*100,digits=1),
            pct_home_student = round_half_up(sum(home_student_y_n=="y",na.rm=T)/total_n*100,digits=1))
```

```{r }
overtime_data %>%
  write_rds(here(path="Data",
                 "2020_10_31_overtime_data.rds"))
```
